
#In this file, we will illustrate how to get a good fit of the conditional intensities (if there is time dependent covariates)/hazards (if no time dependent covariates) and then use the g computation formula to get a plug in estimate. 
#We will illustrate this in discrete time. The continuous time will be briefly mentioned if time is allowed.

#Load a dataset for illustration
```{r}
library(MASS)
library(ltmle)
library(survival)

data <- bladder1

data <- data %>% 
  dplyr::select(-c(rtumor,rsize,enum,recur))

#redefine treatment to numeric
#0 = placebo, 1 = pyridoxine, 2 = thiotepa
data$treatment <- as.numeric(data$treatment) - 1

#for now, only use the first recurrence for a patient
data <- data %>% filter(start == 0) %>%
  mutate(time = stop) %>%
  dplyr::select(-c(start,stop,id))

full_data <- data

#make 3 different datasets for 3 levels of comparison
data01 <- data %>% filter(treatment %in% c(0,1))
data02 <- data %>% filter(treatment %in% c(0,2))
data12 <- data %>% filter(treatment %in% c(1,2))

data0 <- data %>% filter(treatment == 0)

```

#Prepare our data in a discrete time scale and one censoring

#Time scale
```{r}
#ltmle only handle discrete time. Here, time is in days, maybe too detailed and thus in order to pull across observations, we need a less detailed version, say years

#Before we continue our discretization, let's visualize the time collected in days.

# hist(data$time)
# data$time<-as.integer(data$time/3)+1
hist(data$time)
max_time<-max(data$time)
```

#Censoring

```{r}
library(dplyr)
#next change status = 3 to 0
#assume death is type of censoring
# for now change status = 2 to 0 as well... 0 censored, 1 recurrence
data$status[data$status == 3 | data$status == 2] <- 0
```

#Add a id column and created a repeated data format.
```{r}
n<-dim(data)[1]
data$id<-c(1:n)
```

#Suppose Censoring happen before failure at any time point
#Info for failure across time
```{r}

failure_repeated_data<-NULL

for(i in 1:n){
  #if it is a censoring at time t, then we know the failure hasn't happened before time t
  if (data$time[i]>1 & data$status[i]==0){
    time_col<-c(1:(data$time[i]-1))
    count_col<-rep(0,data$time[i]-1)
    baseline_part<-data[i,c(1,2,3,6)] %>% slice(rep(1:n(), each = data$time[i]-1))
    failure_data_i<-cbind(time_col,count_col,baseline_part)
    failure_repeated_data <-rbind(failure_repeated_data,failure_data_i)               
  }
  #if it is a failure at time t, then we know the failure hasn't happened before time t and happen at time t
  if (data$status[i]==1){
    time_col<-c(1:(data$time[i]))
    count_col<-c(rep(0,data$time[i]-1),1)
    baseline_part<-data[i,c(1,2,3,6)] %>% slice(rep(1:n(), each = data$time[i]))
    failure_data_i<-cbind(time_col,count_col,baseline_part)
    failure_repeated_data <-rbind(failure_repeated_data,failure_data_i)  
  }
}
```

#Fit hazard/intensity model of interest using super learner
```{r}
library(SuperLearner)
sl = SuperLearner(Y = failure_repeated_data$count_col, X =failure_repeated_data[,c(1,3,4,5)] , family = binomial(),
  SL.library = c("SL.mean", "SL.glmnet", "SL.ranger"))
sl
```
#Get the g computation evalutation trhough monte carlo simulation (suppose interested in A=1 counterfactual world)
```{r}
sim_one<-function(seed){
  failed<-0
  baseline_part<-data[ sample(1:n, 1, replace=TRUE),c(2,3)]
  baseline_treatment_part<-cbind(baseline_part,treatment=1)
  time_fail<-1
  while(failed==0 &time_fail<=max_time){
    xinput<-cbind(time_col=time_fail,baseline_treatment_part)
    # onlySL is set to TRUE so we don't fit algorithms that had weight = 0, saving computation.
    pred = predict(sl, xinput, onlySL = TRUE)
    failed<-rbinom(1,1,as.numeric(pred[1]))
    if(failed==0){
      time_fail<-time_fail+1
    }
  }
  return(time_fail)
}

```

#In our example, max_time=16, so 17 means not failed at time 16
```{r}
library(doParallel)
sim_n<-200
num_cores <- detectCores()-2
registerDoParallel(cores = num_cores)

sim_result<-foreach(i = 1:sim_n, .combine = rbind) %dopar% {
                return(sim_one(i))
}

result_vector<-as.numeric(sim_result)
```

#If interested in P(T>3), P(T>5), P(T>16)
```{r}
#P(T>6)
mean(result_vector>6)

#P(T>12)
mean(result_vector>12)

#P(T>24)
mean(result_vector>24)
```



#Inference will be based on bootstrap, illustrated in another file



