---
title: "LTMLE for survtmle"
author: "Nolan Gunter"
date: "11/14/2023"
output: pdf_document
---

This document is for replicating Yi's code for our survival project, with the bladder1 dataset.

#Load Bladder dataset

```{r}
library(survival)
library(ltmle)
library(ggplot2)
library(mvtnorm)
library(dplyr)
library(purrr)
library(rlang)

data <- bladder1

data <- data %>% 
  select(-c(rtumor,rsize,enum,recur))

#next change status = 3 to 0
#assume death is type of censoring
# for now change status = 2 to 0 as well...
data$status[data$status == 3 | data$status == 2] <- 0

#redefine treatment to numeric
#0 = placebo, 1 = pyridoxine, 2 = thiotepa
data$treatment <- as.numeric(data$treatment) - 1

#for now, only use the first recurrence for a patient
data <- data %>% filter(start == 0) %>%
  mutate(time = stop) %>%
  select(-c(start,stop,id))

full_data <- data
```

Use this block of code ONLY if you are using the simulated data

```{r}
data <- sim
```

```{r}
#make 3 different datasets for 3 levels of comparison
data01 <- data %>% filter(treatment %in% c(0,1))
data02 <- data %>% filter(treatment %in% c(0,2))
data12 <- data %>% filter(treatment %in% c(1,2))

#data <- data01

data <- data02
data$treatment[data$treatment == 2] <- 1

#data <- data12
#data$treatment <- data$treatment - 1

#time survival time in month, possibly censored.
#status 0 censored, 1 recurrence, 2 death from bladder disease, 3 dead from other causes
```

Some facts about using ltmle:

The estimates returned by ltmle are of a treatment specific mean, E[Ya], the mean of the final treatment node, where all treatment nodes, A, are set to a  and all censoring nodes C are set to 1 (uncensored).

data should be a data frame where the order of the columns corresponds to the time-ordering of the model.

*In censoring columns (Cnodes): factor with two levels: "censored" and "uncensored". Thehelper function BinaryToCensoring can be used to create these factors

*In treatment columns (Anodes): 1 = treated, 0 = untreated (must be binary)

*In event columns (Ynodes): If survivalOutcome is TRUE, then Y nodes are treated as indicators of a one-time event

*time-dependent covariate columns (Lnodes): can be any numeric data

*Data in Cnodes, Anodes, Lnodes and Ynodes are not used after (to the right of) censoring (or an event when survivalOutcome==TRUE) and may be coded as NA or any other value.

*Columns in data that are before (to the left of) the first of Cnodes or Anodes are treated as baseline variables, even if they are specified as Lnodes.

*After the first of Cnodes, Anodes, Ynodes, or Lnodes, every column must be in one of Cnodes, Anodes, Ynodes, or Lnodes.

*If survivalOutcome is TRUE, all Y values are indicators of an event (e.g. death) at or before the current time, where 1 = event and 0 = no event. The events in Ynodes must be of the form where once Y jumps to 1, Y remains 1 at subsequent nodes.

#Prepare our data for ltmle usage

#Time scale
```{r}
#ltmle only handle discrete time
#Before we continue our discretization, let's visualize the time collected in days.
hist(data$time)
```

#Now, let's create a dataset that keeps track of the time-dependent status in each row. 

```{r}
#For each row in my original data, I know the event happen at the recorded time, no matter it is a censoring event or a failure event. This implies that before that time, the Cnodes are 0 (no censroing) and the Ynodes is 0 (not failed yet).

#At each time point, the Cnode comes before Ynode. Since If not censored at that time, we can observe whether failure happened at that time or not.

#Note that the time 0 treatment is already after the baseline nodes, so don't need to move it to the position after the baseline columns.

max_time <- max(data$time)

c_column_name <- paste0("C_", 0:max(data$time))
y_column_name <- paste0("Y_", 0:max(data$time))

for(i in 0:max(data$time)){
  data <- data %>% 
    mutate(!!c_column_name[i+1] := ifelse(time==i & status==1,1,ifelse(time<i,ifelse(status==0,0,NA),0)),
           !!y_column_name[i+1] := ifelse(time<=i & status==0,1,ifelse(time<=i,NA,0)))
  
}

data <- data %>% dplyr::select(-time,-status)

#We will use the BinaryToCensoring function to help us changing the Cnodes to factor object to be used in ltmle
for(j in 0:max_time){
  data[,c_column_name[j+1]] <- BinaryToCensoring(is.censored=data[,c_column_name[j+1]])
}
```

#Fit ltmle

```{r}
#Now, we need to tell the ltmle what are the A nodes, L nodes, C nodes and Y nodes
time <- 24
Anodes <- c("treatment") #can give by column name
Lnodes<- c(2,3) #can give by column number
Cnodes<- c_column_name[1:(time+1)]
Ynodes<- y_column_name[1:(time+1)]
```

#Set up a superLearner library to use
```{r}
library(SuperLearner)
library(arm) #the SL.bayesglm needs this package
#You can try to put more complicated superLearner libraries
SL.library <- c("SL.glm", "SL.bayesglm", "SL.mean")

tmp_dat <- data %>% dplyr::select(treatment,number,size,Cnodes,Ynodes)
  
results.ltmle <- ltmle(
  data = tmp_dat,
  survivalOutcome = TRUE,
  Anodes = Anodes, Lnodes = Lnodes, Cnodes = Cnodes,  Ynodes = Ynodes,
  abar = 1, #to specify the intervention of interest to 1
  variance.method = "ic",
  SL.library=SL.library
)

#This gives you the tmle estimate for the counterfactual P(T1<=time) in a world where treatment is set to 1 and there is no censoring.
#Since this gives you the  E(Y_5) under no censoring and intervention 1. Due to Y_time=1 or 0, this is the same as P(Y_time=1), but Y_time=1 means T<=time.
results.ltmle$estimates[1]
#This gives you a vector of Influence curve evaluation for P(T1<=time) at each observation O_i, so that you can use it for simultaneous interval construction.

#results.ltmle$IC$tmle

c(results.ltmle$estimates[1] - 1.96*(sd(results.ltmle$IC$tmle))/sqrt(nrow(tmp_dat)),
  results.ltmle$estimates[1] + 1.96*(sd(results.ltmle$IC$tmle))/sqrt(nrow(tmp_dat)))

###
#Now if one is interested in the 1-P(T1>time)
#tmle estimate for 1-P(T1>time)
1-results.ltmle$estimates[1]
#IC(O_i) for 1-P(T1>time)

#-results.ltmle$IC$tmle
```

```{r}
### THIS IS WHAT I USED FOR INITIAL LTMLE RESULTS

#Now if one is interested in the P(T1<=TIME)-P(T0<=TIME)
#tmle estimate for P(T1<=TIME)-P(T0<=TIME)
results.ltmle$estimates[1]-results.ltmle_0$estimates[1]

#IC(O_i) for P(T1<=time)-P(T0<=time)
results.ltmle$IC$tmle -results.ltmle_0$IC$tmle 

ci <- c(results.ltmle$estimates[1]-results.ltmle_0$estimates[1] - 1.96*(sd(results.ltmle$IC$tmle -results.ltmle_0$IC$tmle))/sqrt(nrow(data_lt)),
        results.ltmle$estimates[1]-results.ltmle_0$estimates[1] + 1.96*(sd(results.ltmle$IC$tmle -results.ltmle_0$IC$tmle))/sqrt(nrow(data_lt)))
```


#Now one has everything for mutliple time points, then one can go over the Simultaneous_CI.Rmd files row 141-182 to construct simultaneous CI.

```{r}
### replicating Yi's code for our data

###Use this chunk if you are using simulated data
failure1_data <- sim

###Use this chunk if you are using real data
failure1_data <- full_data



Ttilde <- failure1_data$time
Delta <- failure1_data$status == 1
n <- length(Ttilde)

# do a KM fit
## EDIT: Yi regressed against 1 instead of treatment
#km <- survfit(Surv(Ttilde, Delta, type = "right") ~ failure1_data$treatment, type = "kaplan-meier", 
#              conf.int = .95)

#this part will estimate overall survival
km <- survfit(Surv(Ttilde, Delta, type = "right") ~ 1, type = "kaplan-meier", 
              conf.int = .95)

# make step functions
km_step <- stepfun(km$time, c(1, km$surv))
km_lower <- stepfun(km$time, c(1, km$lower))
km_upper <- stepfun(km$time, c(1, km$upper))

range <- seq(0:max(failure1_data$time))

km_data <- data.frame(time = range, 
                      lower = km_lower(range),
                      KM = km_step(range),
                      upper = km_upper(range)
                      )

km_data_long <- data.frame(time = c(range,range,range),
                           estimate = c(km_data$lower,km_data$KM,km_data$upper),
                           label = c(rep("lower",length(range)),
                                     rep("KM",length(range)),
                                     rep("upper",length(range))))


ggplot(data = km_data_long, aes(x = time, y = estimate, color = label)) + 
  geom_point() + geom_line() + labs(y = "Estimated Probability", x = "Time (months)") + theme_minimal()
```

```{r}
kmcurve=survfit(Surv(failure1_data$time,failure1_data$status)~failure1_data$treatment)
survdiff(Surv(failure1_data$time,failure1_data$status)~failure1_data$treatment)
survminer::ggsurvplot(kmcurve,data=failure1_data)
```


```{r}
#STEP1 start
# order observed data according to increasing order Ttilde
T_ord <- Ttilde[order(Ttilde)]
D_ord <- Delta[order(Ttilde)]

# all known death times in order
T_death <- T_ord[D_ord]
    
# compute the hazard (lambda) at all known death times
nrisk <- sapply(T_death, function(x) sum(Ttilde >= x))
lambda <- 1/nrisk

# compute the prob of surviving up to or past each death time
Pbar <- sapply(T_death, function(x) mean(Ttilde >= x))
    
calculate_IC <- function(Ttilde, Delta, time, n) {
    sum((-(T_death <= time)*km_step(time)/(1 - lambda))*
          lambda^2*(1 - lambda)^2/((1 - lambda)^2/n + lambda^2*Pbar)*
          ((Ttilde == T_death & Delta == 1)/lambda - (Ttilde > T_death)/(1 - lambda)),na.rm=T)
}

# compute the IC matrix for times 40, 50, 60 where each column is an IC
times <- c(12,24,36)
IC_matrix <- sapply(times, function(t) {
    unlist(lapply(1:n, function(i) calculate_IC(Ttilde[i], Delta[i], t, n)))
})
#STEP1 end
    
#STEP2 start
# compute the correlation of the IC matrix
Sigma <- cor(IC_matrix)
# randomly draw from a multivariate Normal(0,Sigma) distribution
mean <- rep(0, length(times))
z <- rmvnorm(1e6, mean, Sigma)
    
# identify row-wise maximum of the absolute value of each MVN value in z
z_abs <- apply(z, 1, function(x) max(abs(x))) 
# identify 95th quantile of z_abs, which is the se to use for simultaneous CI
z_95 <- quantile(z_abs, .95)
# note how the se_simultaneous is larger when demanding simultaneous coverage
# of the truth, and it directly follows that simultaneous CIs are wider.
    
# check if all of the simultaneous CI's covered the truth
sd_IC1 <- sd(IC_matrix[,1])/sqrt(n)
sd_IC2 <- sd(IC_matrix[,2])/sqrt(n)
sd_IC3 <- sd(IC_matrix[,3])/sqrt(n)
    
ci1 <- c(km_step(times[1]) - z_95*sd_IC1, km_step(times[1]) + z_95*sd_IC1)
    
ci2 <- c(km_step(times[2]) - z_95*sd_IC2, km_step(times[2]) + z_95*sd_IC2)
    
ci3 <- c(km_step(times[3]) - z_95*sd_IC3, km_step(times[3]) + z_95*sd_IC3)
    
ci1
ci2
ci3
```

