---
title: "LTMLE for survtmle"
author: "Nolan Gunter"
date: "11/14/2023"
output: pdf_document
---

This document is for replicating Yi's code for our survival project, with the bladder1 dataset.

#Load Bladder dataset

```{r}
library(survival)
library(ltmle)
library(ggplot2)
library(mvtnorm)
library(dplyr)
library(purrr)
library(rlang)

data <- bladder1

data <- data %>% 
  dplyr::select(-c(rtumor,rsize,enum,recur))

#next change status = 3 to 0
#assume death is type of censoring
# for now change status = 2 to 0 as well...
data$status[data$status == 3 | data$status == 2] <- 0

#redefine treatment to numeric
#0 = placebo, 1 = pyridoxine, 2 = thiotepa
data$treatment <- as.numeric(data$treatment) - 1

#for now, only use the first recurrence for a patient
data <- data %>% filter(start == 0) %>%
  mutate(time = stop) %>%
  dplyr::select(-c(start,stop,id))

full_data <- data
```

Use this block of code ONLY if you are using the simulated data

```{r}

```

```{r}
#make 3 different datasets for 3 levels of comparison
data01 <- data %>% filter(treatment %in% c(0,1))
data02 <- data %>% filter(treatment %in% c(0,2))
data12 <- data %>% filter(treatment %in% c(1,2))

#data <- data01

data <- data02
data$treatment[data$treatment == 2] <- 1

#data <- data12
#data$treatment <- data$treatment - 1

#time survival time in month, possibly censored.
#status 0 censored, 1 recurrence, 2 death from bladder disease, 3 dead from other causes
```

Some facts about using ltmle:

The estimates returned by ltmle are of a treatment specific mean, E[Ya], the mean of the final treatment node, where all treatment nodes, A, are set to a  and all censoring nodes C are set to 1 (uncensored).

data should be a data frame where the order of the columns corresponds to the time-ordering of the model.

*In censoring columns (Cnodes): factor with two levels: "censored" and "uncensored". Thehelper function BinaryToCensoring can be used to create these factors

*In treatment columns (Anodes): 1 = treated, 0 = untreated (must be binary)

*In event columns (Ynodes): If survivalOutcome is TRUE, then Y nodes are treated as indicators of a one-time event

*time-dependent covariate columns (Lnodes): can be any numeric data

*Data in Cnodes, Anodes, Lnodes and Ynodes are not used after (to the right of) censoring (or an event when survivalOutcome==TRUE) and may be coded as NA or any other value.

*Columns in data that are before (to the left of) the first of Cnodes or Anodes are treated as baseline variables, even if they are specified as Lnodes.

*After the first of Cnodes, Anodes, Ynodes, or Lnodes, every column must be in one of Cnodes, Anodes, Ynodes, or Lnodes.

*If survivalOutcome is TRUE, all Y values are indicators of an event (e.g. death) at or before the current time, where 1 = event and 0 = no event. The events in Ynodes must be of the form where once Y jumps to 1, Y remains 1 at subsequent nodes.

#Prepare our data for ltmle usage

#Time scale
```{r}
#ltmle only handle discrete time
#Before we continue our discretization, let's visualize the time collected in days.
hist(data$time)
```

#Now, let's create a dataset that keeps track of the time-dependent status in each row. 

```{r}
#For each row in my original data, I know the event happen at the recorded time, no matter it is a censoring event or a failure event. This implies that before that time, the Cnodes are 0 (no censroing) and the Ynodes is 0 (not failed yet).

#At each time point, the Cnode comes before Ynode. Since If not censored at that time, we can observe whether failure happened at that time or not.

#Note that the time 0 treatment is already after the baseline nodes, so don't need to move it to the position after the baseline columns.

max_time <- max(data$time)

c_column_name <- paste0("C_", 0:max(data$time))
y_column_name <- paste0("Y_", 0:max(data$time))

for(i in 0:max(data$time)){
  data <- data %>% 
    mutate(!!c_column_name[i+1] := ifelse(time==i & status==1,1,ifelse(time<i,ifelse(status==0,0,NA),0)),
           !!y_column_name[i+1] := ifelse(time<=i & status==0,1,ifelse(time<=i,NA,0)))
  
}

data <- data %>% dplyr::select(-time,-status)

#We will use the BinaryToCensoring function to help us changing the Cnodes to factor object to be used in ltmle
for(j in 0:max_time){
  data[,c_column_name[j+1]] <- BinaryToCensoring(is.censored=data[,c_column_name[j+1]])
}
```


#Fit ltmle example

```{r} 
#Now, we need to tell the ltmle what are the A nodes, L nodes, C nodes and Y nodes
time <- 24
Anodes <- c("treatment") #can give by column name
Lnodes<- c(2,3) #can give by column number
Cnodes<- c_column_name[1:(time+1)]
Ynodes<- y_column_name[1:(time+1)]


library(SuperLearner)
library(arm) #the SL.bayesglm needs this package
#You can try to put more complicated superLearner libraries
SL.library <- c("SL.glm", "SL.bayesglm", "SL.mean")

tmp_dat <- data %>% dplyr::select(treatment,number,size,Cnodes,Ynodes)
  
results.ltmle <- ltmle(
  data = tmp_dat,
  survivalOutcome = TRUE,
  Anodes = Anodes, Lnodes = Lnodes, Cnodes = Cnodes,  Ynodes = Ynodes,
  abar = 1, #to specify the intervention of interest to 1
  variance.method = "ic",
  SL.library=SL.library
)

#This gives you the tmle estimate for the counterfactual P(T1<=time) in a world where treatment is set to 1 and there is no censoring.
#Since this gives you the  E(Y_5) under no censoring and intervention 1. Due to Y_time=1 or 0, this is the same as P(Y_time=1), but Y_time=1 means T<=time.
results.ltmle$estimates[1]
#This gives you a vector of Influence curve evaluation for P(T1<=time) at each observation O_i, so that you can use it for simultaneous interval construction.

results.ltmle$IC$tmle

c(results.ltmle$estimates[1] - 1.96*(sd(results.ltmle$IC$tmle))/sqrt(nrow(tmp_dat)),
  results.ltmle$estimates[1] + 1.96*(sd(results.ltmle$IC$tmle))/sqrt(nrow(tmp_dat)))

###
#Now if one is interested in the 1-P(T1>time)
#tmle estimate for 1-P(T1>time)
1-results.ltmle$estimates[1]
#IC(O_i) for 1-P(T1>time)

#-results.ltmle$IC$tmle
```


#Fit ltmle for t = 24

```{r} 
#Now, we need to tell the ltmle what are the A nodes, L nodes, C nodes and Y nodes
time <- 24
Anodes <- c("treatment") #can give by column name
Lnodes<- c(2,3) #can give by column number
Cnodes<- c_column_name[1:(time+1)]
Ynodes<- y_column_name[1:(time+1)]


library(SuperLearner)
library(arm) #the SL.bayesglm needs this package
#You can try to put more complicated superLearner libraries
SL.library <- c("SL.glm", "SL.bayesglm", "SL.mean")

tmp_dat <- data %>% dplyr::select(treatment,number,size,Cnodes,Ynodes)
  
results.ltmle.24 <- ltmle(
  data = tmp_dat,
  survivalOutcome = TRUE,
  Anodes = Anodes, Lnodes = Lnodes, Cnodes = Cnodes,  Ynodes = Ynodes,
  abar = 1, #to specify the intervention of interest to 1
  variance.method = "ic",
  SL.library=SL.library
)

#This gives you the tmle estimate for the counterfactual P(T1<=time) in a world where treatment is set to 1 and there is no censoring.
#Since this gives you the  E(Y_5) under no censoring and intervention 1. Due to Y_time=1 or 0, this is the same as P(Y_time=1), but Y_time=1 means T<=time.
results.ltmle.24$estimates[1]
#This gives you a vector of Influence curve evaluation for P(T1<=time) at each observation O_i, so that you can use it for simultaneous interval construction.

results.ltmle.24$IC$tmle

c(results.ltmle.24$estimates[1] - 1.96*(sd(results.ltmle.36$IC$tmle))/sqrt(nrow(tmp_dat)),
  results.ltmle.24$estimates[1] + 1.96*(sd(results.ltmle.36$IC$tmle))/sqrt(nrow(tmp_dat)))

###
#Now if one is interested in the 1-P(T1>time)
#tmle estimate for 1-P(T1>time)
1-results.ltmle.24$estimates[1]
#IC(O_i) for 1-P(T1>time)

#-results.ltmle$IC$tmle
```



#Fit ltmle for t = 36

```{r} 
#Now, we need to tell the ltmle what are the A nodes, L nodes, C nodes and Y nodes
time <- 36
Anodes <- c("treatment") #can give by column name
Lnodes<- c(2,3) #can give by column number
Cnodes<- c_column_name[1:(time+1)]
Ynodes<- y_column_name[1:(time+1)]


library(SuperLearner)
library(arm) #the SL.bayesglm needs this package
#You can try to put more complicated superLearner libraries
SL.library <- c("SL.glm", "SL.bayesglm", "SL.mean")

tmp_dat <- data %>% dplyr::select(treatment,number,size,Cnodes,Ynodes)
  
results.ltmle.36 <- ltmle(
  data = tmp_dat,
  survivalOutcome = TRUE,
  Anodes = Anodes, Lnodes = Lnodes, Cnodes = Cnodes,  Ynodes = Ynodes,
  abar = 1, #to specify the intervention of interest to 1
  variance.method = "ic",
  SL.library=SL.library
)

#This gives you the tmle estimate for the counterfactual P(T1<=time) in a world where treatment is set to 1 and there is no censoring.
#Since this gives you the  E(Y_5) under no censoring and intervention 1. Due to Y_time=1 or 0, this is the same as P(Y_time=1), but Y_time=1 means T<=time.
results.ltmle.36$estimates[1]
#This gives you a vector of Influence curve evaluation for P(T1<=time) at each observation O_i, so that you can use it for simultaneous interval construction.

results.ltmle.36$IC$tmle

c(results.ltmle.36$estimates[1] - 1.96*(sd(results.ltmle.36$IC$tmle))/sqrt(nrow(tmp_dat)),
  results.ltmle.36$estimates[1] + 1.96*(sd(results.ltmle.36$IC$tmle))/sqrt(nrow(tmp_dat)))

###
#Now if one is interested in the 1-P(T1>time)
#tmle estimate for 1-P(T1>time)
1-results.ltmle.36$estimates[1]
#IC(O_i) for 1-P(T1>time)

#-results.ltmle$IC$tmle
```


#Fit ltmle for t = 48

```{r} 
#Now, we need to tell the ltmle what are the A nodes, L nodes, C nodes and Y nodes
time <- 48
Anodes <- c("treatment") #can give by column name
Lnodes<- c(2,3) #can give by column number
Cnodes<- c_column_name[1:(time+1)]
Ynodes<- y_column_name[1:(time+1)]


library(SuperLearner)
library(arm) #the SL.bayesglm needs this package
#You can try to put more complicated superLearner libraries
SL.library <- c("SL.glm", "SL.bayesglm", "SL.mean")

tmp_dat <- data %>% dplyr::select(treatment,number,size,Cnodes,Ynodes)
  
results.ltmle.48 <- ltmle(
  data = tmp_dat,
  survivalOutcome = TRUE,
  Anodes = Anodes, Lnodes = Lnodes, Cnodes = Cnodes,  Ynodes = Ynodes,
  abar = 1, #to specify the intervention of interest to 1
  variance.method = "ic",
  SL.library=SL.library
)

#This gives you the tmle estimate for the counterfactual P(T1<=time) in a world where treatment is set to 1 and there is no censoring.
#Since this gives you the  E(Y_5) under no censoring and intervention 1. Due to Y_time=1 or 0, this is the same as P(Y_time=1), but Y_time=1 means T<=time.
results.ltmle.48$estimates[1]
#This gives you a vector of Influence curve evaluation for P(T1<=time) at each observation O_i, so that you can use it for simultaneous interval construction.

results.ltmle.48$IC$tmle

c(results.ltmle.48$estimates[1] - 1.96*(sd(results.ltmle.48$IC$tmle))/sqrt(nrow(tmp_dat)),
  results.ltmle.48$estimates[1] + 1.96*(sd(results.ltmle.48$IC$tmle))/sqrt(nrow(tmp_dat)))

###
#Now if one is interested in the 1-P(T1>time)
#tmle estimate for 1-P(T1>time)
1-results.ltmle.48$estimates[1]
#IC(O_i) for 1-P(T1>time)

#-results.ltmle$IC$tmle
```




#Now one has everything for mutliple time points, then one can go over the Simultaneous_CI.Rmd files row 141-182 to construct simultaneous CI.



```{r}

n <- length(results.ltmle.24$IC$tmle)
IC_matrix <- cbind(results.ltmle.24$IC$tmle, results.ltmle.36$IC$tmle, results.ltmle.48$IC$tmle)
#STEP2 start
# compute the correlation of the IC matrix
Sigma <- cor(IC_matrix)
# randomly draw from a multivariate Normal(0,Sigma) distribution
mean <- rep(0, length(times))
z <- rmvnorm(1e6, mean, Sigma)
    
# identify row-wise maximum of the absolute value of each MVN value in z
z_abs <- apply(z, 1, function(x) max(abs(x))) 
# identify 95th quantile of z_abs, which is the se to use for simultaneous CI
z_95 <- quantile(z_abs, .95)
# note how the se_simultaneous is larger when demanding simultaneous coverage
# of the truth, and it directly follows that simultaneous CIs are wider.
    
# check if all of the simultaneous CI's covered the truth
sd_IC.24 <- sd(IC_matrix[,1])/sqrt(n)
sd_IC.36 <- sd(IC_matrix[,2])/sqrt(n)
sd_IC.48 <- sd(IC_matrix[,3])/sqrt(n)

ltmle.est.24 <- results.ltmle.24$estimates[1] 
ltmle.est.36 <- results.ltmle.36$estimates[1] 
ltmle.est.48 <- results.ltmle.48$estimates[1]    

ci.24 <- c(ltmle.est.24-z_95*sd_IC.24, ltmle.est.24+z_95*sd_IC.24)
    
ci.36 <- c(ltmle.est.36-z_95*sd_IC.36, ltmle.est.36+z_95*sd_IC.36)
    
ci.48 <- c(ltmle.est.48-z_95*sd_IC.48, ltmle.est.48+z_95*sd_IC.48)
    
ci.24
ci.36
ci.48

ltmle.est.24
ltmle.est.36
ltmle.est.48

ci.pt.24 <- c(results.ltmle.24$estimates[1] - 1.96*(sd(results.ltmle.24$IC$tmle))/sqrt(nrow(tmp_dat)),
  results.ltmle.24$estimates[1] + 1.96*(sd(results.ltmle.24$IC$tmle))/sqrt(nrow(tmp_dat)))

ci.pt.36 <- c(results.ltmle.36$estimates[1] - 1.96*(sd(results.ltmle.36$IC$tmle))/sqrt(nrow(tmp_dat)),
  results.ltmle.36$estimates[1] + 1.96*(sd(results.ltmle.36$IC$tmle))/sqrt(nrow(tmp_dat)))

ci.pt.48 <- c(results.ltmle.48$estimates[1] - 1.96*(sd(results.ltmle.48$IC$tmle))/sqrt(nrow(tmp_dat)),
  results.ltmle.48$estimates[1] + 1.96*(sd(results.ltmle.48$IC$tmle))/sqrt(nrow(tmp_dat)))

ci.pt.24
ci.pt.36
ci.pt.48

```

